Flowchart of Current Algorithm

### 2009-04-23 docs/flowchart/flowchart.tex
more progress on spherical base state

\section{Changes Between Paper 3 and Paper 4}
\begin{enumerate}
\item {\tt fill\_3d} and {\tt average} algorithms.
\item We update temperature after each {\bf React State} as well as the advection step.
\item We have created a {\tt burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
\item Use corner coupling in advection.
\item We use {\tt use\_tfromp} to update temperature using $T=T(\rho,X_k,p_0)$ rather 
than $T=T(\rho,h,X_k)$.
\end{enumerate}
Changes which modify enthalpy update.  These are separate because in this paper,
enthalpy is only used as a diagnostic since we use {\tt use\_tfromp}.  If we don't
mention these items in paper IV, they should me noted in paper 4.5.
\begin{enumerate}
\item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
update to $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
\item In paper III we discretized the enthalpy evolution equation in
full, non-perturbation form, $\rho h$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
This is more like paper II.
\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
\end{enumerate}
\section{Changes Between Papers 4 and Papers 4.5}
\begin{enumerate}
\item We split the update of the density and enthalpy in steps 4 and 8.
Now, density is updated first, allowing us to compute time-centered values of 
$\psi$ that are then used in the enthalpy update.
\item We split {\bf Advect Base} into separate {\bf Advect Base Density},
{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  The calling sequence
is in the main flow chart.
\item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
plane-parallel and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical,
eliminating the need for $\nabla\cdot\etarho$.  {\bf Correct Base} is now used
analogously for setting $(\rho h)_0$ after {\bf React State}.
\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
in spherical.  This means that {\bf React Base} no longer exists, and the intermediate 
states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
{\bf Diffuse Base}.
\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
occurs after the full state density update and  before {\bf Advect Base Pressure}.  
Since the full enthalpy advance ``feels'' the effects of the time-centered $\etarho$, 
we should rely on a $\rho_0^{n+1}$ that also feels the effects.
\end{enumerate}
Changes specific to spherical:
\begin{enumerate}
\item Discuss {\bf Advect Base} routines and {\bf Correct Base} routine.
\item $\etarho$ and $\psi$ equations.
\end{enumerate}
\section{Changes Between Papers 4.5 and Paper 5}
\begin{enumerate}
\item Use PPM limiters in advection.
\end{enumerate}
\section{Changes Between Paper 5 and XRB Paper}
\begin{enumerate}
\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
\item We added the volume-discrepancy term to the velocity constraint equation,
controlled by the input parameter, {\tt dpdt\_factor}.
\end{enumerate}

@@

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite VODE).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
(paper III; eq.~[8])
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We update temperature using
$T^{\outp} = T(\rho^\outp,h^\outp,X_k^\outp)$ if {\tt use\_tfromp = FALSE} or 
$T^{\outp} = T(\rho^\outp,p_0,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
Note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

@@

@@

\item[plane-parallel:] We discretize the base state density evolution 
equation (paper III; eq.~[29]) without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ]
         \nonumber \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are found via the
procedure described in paper II, Appendix A.\\

@@

First we compute a cell-centered $\psi^\outp$ using
\begin{equation}
 \psi_j^\outp = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
\end{equation}
Then, we discretize the base state pressure evolution equation (paper III; eq.~[31]),
\begin{equation}
\frac{\partial p_0}{\partial t} = 
-w_0 \frac{\partial p_0}{\partial r} + \psi \enskip ,
\label{eq:p0upd_new}
\end{equation}
to compute the new base state pressure,
\begin{equation}
p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\outp 
\enskip,
\end{equation}
where the interface states are again found via the procedure described
in paper II, Appendix A.\\

@@

\item At the beginning of each time step, we need an estimate for the time-centered
source term in the velocity divergence constraint,
\begin{equation}
  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext 
  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip ,
\label{eq:defineS} 
\end{equation}
(paper III; eq.~[19]).

@@

\item Construct $w_0^{\nph,\star}$

\item[plane-parallel:] We integrate
\begin{equation}
\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g 
\enskip \label{eq:divw0} ,
\end{equation}
using the lagged $\psi^{\nmh}$ and the volume discrepancy correction with $f=0.3$,
\begin{equation}
\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S^{\nph,\star}}
- \frac{1}{\gammabar^{n} p_0^{n}}
\left\{ \psi^{\nmh} + 
f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
\enskip .
\end{equation}

\item[spherical:] See Section \ref{Sec:Spherical w0 Calculation} for working notes.

@@

The local velocity field is described by
\begin{equation}
\frac{\partial\Ubt}{\partial t} = - \Ubt \cdotb \nablab \Ubt - w_0 \frac{\partial \Ubt}{\partial r}
                                 - \left(\Ubt \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                 - \frac{1}{\rho} \nablab\pi
                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
 ,
\end{equation}
(paper III; eq.~[37]).

>>>>

\section{Changes Between Paper 3 and Paper 4}
\begin{enumerate}
\item {\tt fill\_3d} and {\tt average} algorithms.
\item We update temperature after each {\bf React State} as well as the advection step.
\item We have created a {\tt burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
\item Use corner coupling in advection.
\item We use {\tt use\_tfromp} to update temperature using $T=T(\rho,X_k,p_0)$ rather 
than $T=T(\rho,h,X_k)$.
\end{enumerate}
Changes which modify enthalpy update.  These are separate because in this paper,
enthalpy is only used as a diagnostic since we use {\tt use\_tfromp}.  If we don't
mention these items in paper IV, they should me noted in paper 4.5.
\begin{enumerate}
\item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
update to $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
\item In paper III we discretized the enthalpy evolution equation in
full, non-perturbation form, $\rho h$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
This is more like paper II.
\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
\end{enumerate}
\section{Changes Between Papers 4 and Papers 4.5}
\begin{enumerate}
\item We split the update of the density and enthalpy in steps 4 and 8.
Now, density is updated first, allowing us to compute time-centered values of 
$\psi$ that are then used in the enthalpy update.
\item We split {\bf Advect Base} into separate {\bf Advect Base Density},
{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  The calling sequence
is in the main flow chart.
\item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
plane-parallel and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical,
eliminating the need for $\nabla\cdot\etarho$.  {\bf Correct Base} is now used
analogously for setting $(\rho h)_0$ after {\bf React State}.
\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
in spherical.  This means that {\bf React Base} no longer exists, and the intermediate 
states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
{\bf Diffuse Base}.
\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
occurs after the full state density update and  before {\bf Advect Base Pressure}.  
Since the full enthalpy advance ``feels'' the effects of the time-centered $\etarho$, 
we should rely on a $\rho_0^{n+1}$ that also feels the effects.
\item {\bf AMR!!!}
\end{enumerate}
Changes specific to spherical:
\begin{enumerate}
\item Discuss {\bf Advect Base} routines and {\bf Correct Base} routine.
\item $\etarho$ and $\psi$ equations.
\end{enumerate}
\section{Changes Between Papers 4.5 and Paper 5}
\begin{enumerate}
\item Use PPM limiters in advection.
\end{enumerate}
\section{Changes Between Paper 5 and XRB Paper}
\begin{enumerate}
\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
\item We added the volume-discrepancy term to the velocity constraint equation,
controlled by the input parameter, {\tt dpdt\_factor}.
\end{enumerate}

@@

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite VODE).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
from equation (8) in paper III,
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We update temperature using
$T^{\outp} = T(\rho^\outp,h^\outp,X_k^\outp)$ if {\tt use\_tfromp = FALSE} or 
$T^{\outp} = T(\rho^\outp,p_0,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
Note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

@@

\item {\bf Compute $w_0$}\\
is the process by which we compute $w_0$.  The arguments are different for plane-parallel
and spherical.
\begin{description}
\item {\bf Compute $w_0$ Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},\psi^{\inp}] 
\rightarrow [w_0^{\outp}]$:

We integrate
\begin{equation}
\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \psi
\enskip \label{eq:divw0} ,
\end{equation}
with the following discretization:
\begin{equation}
\frac{\partial w_0^{\outp}}{\partial r} =  \Sbar^{\inp}
- \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}
\enskip .
\end{equation}
% VOLUME DISCREPANCY STUFF
%\begin{equation}
%\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S^{\nph,\star}}
%- \frac{1}{\gammabar^{n} p_0^{n}}
%\left\{ \psi^{\nmh} + 
%f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
%\enskip .
%\end{equation}

\item {\bf Compute $w_0$ Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},
\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:

We begin with equation (23) in paper III written in spherical coordinates:
\[
\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) =
\beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )\enskip.
\]
We expand the spatial derivative and divide through by $\beta_0$:
\begin{equation}
\label{eq:sphconstraint}
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r} =
 \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t}  \enskip .
\end{equation}
Recalling from paper I that;
\[
\frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}\enskip,
\]
and moving this term to the right hand side,
we can express equation (\ref{eq:sphconstraint}) as:
\begin{eqnarray}
%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
%w_0 \frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} &=&
% \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right ) \nonumber \\
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) &=&
  \Sbar - \frac{1}{\gammabar p_0} \left( 
\frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)\enskip. \label{eq:w01}
\end{eqnarray}
We first solve a piece of this equation for $\ow = w_0 - \dw$:
\begin{eqnarray}
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar \enskip.
\end{eqnarray}
Then we can write
\begin{eqnarray}
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \dw \right ) &=&
  - \frac{1}{\gammabar p_0} \left(
\frac{\partial p_0}{\partial t} + (\ow + \dw)
\frac{\partial p_0}{\partial r} \right)\enskip.\label{eq:w01a}
\end{eqnarray}

Multiplying equation (\ref{eq:w01a}) through by $\gammabar p_0 $,
taking another derivative with respect to $r$, and switching the order
of temporal and spatial derivatives, we get:
\begin{equation}
\label{eq:sphconstraint_2}
\frac{\partial}{\partial r} \left [ \frac{\gammabar p_0 }{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right ]
= -\frac{\partial}{\partial t} \frac{\partial p_0}{\partial r}
  -\frac{\partial}{\partial r} \left ((\ow+\dw) \frac{\partial p_0}{\partial r} \right )\enskip.
\end{equation}

To solve for $\dw$ we will need to substitute for the derivatives of $p_0.$
To do so we start with the equation of hydrostatic equilibrium in a spherical geometry:
\[
\frac{\partial p_0}{\partial r} = -\rho_0 g \, ; \quad
g = \frac{G m_\mathrm{encl}}{r^2}\enskip,
\]
where $m_\mathrm{encl}(r)$ is the mass enclosed at radius $r$ and $G$ is the
gravitational constant.  
Using this, we can then write equation (\ref{eq:sphconstraint_2}) as:
\begin{eqnarray}
\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
 \frac{\partial}{\partial t} \left (\rho_0 g \right )
+\frac{\partial}{\partial r} \left (w_0 \rho_0 g \right) \nonumber \\
&=&
  g \left[ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right]
+ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.\nonumber\\
&&
\label{eq:sphconstraint_3}
\end{eqnarray}
The mass enclosed inside any radius, $r$, is $m_\mathrm{encl}(r) = 4 \pi
\int_0^r \rhozero(s) s^2 ds$, or alternately, $\partial m_\mathrm{encl}/\partial r = 4\pi r^2 \rho_0$.  
The Lagrangian derivative of the enclosed mass is then:
\begin{eqnarray}
\frac{D_0 m_\mathrm{encl}}{D t} &=& \frac{\partial m_\mathrm{encl}}{\partial t} + w_0 \frac{\partial m_\mathrm{encl}}{\partial r} \nonumber \\
 &=& 4\pi \left( \frac{\partial}{\partial t} \int_0^r \rhozero(s) s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
&=& 4\pi \left( \int_0^r \frac{\partial \rho_0}{\partial t} s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
&=& 4\pi \left\{ -\int_0^r \left [ \frac{1}{s^2}\frac{\partial(s^2 \rho_0 w_0)}{\partial s} + \frac{1}{s^2}\frac{\partial (s^2 \etarho)}{\partial s} \right ]s^2 ds + w_0 r^2 \rho_0 \right\} \nonumber \\
&=&  \left . 4\pi \left(-s^2 \rho_0 w_0 \right |_0^r - \left . s^2 \etarho \right |_0^r + w_0 r^2 \rho_0 \right) \nonumber \\
&=& -4\pi r^2 \etarho \label{eq:dmdt}
\end{eqnarray}
where we used the spherical form of equation (29) in paper III, 
\begin{equation}
\label{eq:sph_continuity}
\frac{\partial \rhozero}{\partial t} + \frac{1}{r^2} \frac{\partial (r^2 \rho_0 w_0)}{\partial r} 
+ \frac{1}{r^2} \frac{\partial (r^2 \etarho )}{\partial r}  = 0 \enskip ,
\end{equation}
to eliminate $\partial \rho_0/\partial t$.  We note that in the
absence of any mixing, $\etarho=0$, and  $D_0 m_\mathrm{encl}/Dt = 0.$
%which is another way of saying that the mass inside a fluid element
%does not change during expansion.  
Equation (\ref{eq:dmdt})
allows us to write the Lagrangian change in the gravitational
acceleration as:
\begin{eqnarray}
\frac{D_0 g}{D t} = \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} &=& \frac{D_0}{Dt}\left(\frac{G m_\mathrm{encl}}{r^2}\right) \nonumber \\
&=& G m_\mathrm{encl}\frac{D_0}{D t}\left(\frac{1}{r^2}\right) + \frac{G}{r^2}\frac{D_0 m_\mathrm{encl}}{Dt} \nonumber \\
&=& -\frac{2 w_0 G m_\mathrm{encl}}{r^3} - 4 \pi G \etarho \nonumber \\
&=& -\frac{2 w_0 g}{r} - 4 \pi G \etarho \enskip .
\end{eqnarray}

Putting it all together, equation (\ref{eq:sphconstraint_3}) becomes:
\begin{equation}
 \frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] 
= 
  g \left [ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right ]
+ \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \label{eq:B5} \enskip.
\end{equation}
Finally, we can use equation (\ref{eq:sph_continuity})
to write equation (\ref{eq:B5}) as:
\begin{eqnarray}
\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] && \nonumber \\
&=&
 g \left [ -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 w_0 \rhozero)
           -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 \etarho)
           +              \frac{\partial}{\partial r} (w_0 \rhozero) \right ] + \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \nonumber \\
&=& 
- \frac{g}{r^2} \frac{\partial (r^2 \etarho)}{\partial r} - \frac{4 (\ow + \dw) \rho_0 g}{r} 
- 4 \pi G \rhozero \etarho \label{eq:B7} \enskip .
\end{eqnarray}

\noindent We discretize this elliptic equation in the radial dimension as:
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i} -
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i-1} \right\}
+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{i-\half} \nonumber\\
&=& 
- \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
   - \left[ \frac{4 \ow \rho_0 g}{r} \right]_{i-\half}
   - \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip , \nonumber
\end{eqnarray}
where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
Then, using hydrostatic equilibrium, we expand this to 
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
 \left( \frac{\gammabar p_0}{r^2}\right)_{i  } \frac{[ (r^2 \dw)_{i+\half} - (r^2 \dw)_{i-\half} ]}{\Delta r}
-\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \frac{[ (r^2 \dw)_{i-\half} - (r^2 \dw)_{i-\thalf}]}{\Delta r} \right\} \nonumber \\
&&  - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \dw)_{i-\half}  \nonumber \\
&=&   \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
 - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
  \left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip.
\end{eqnarray}

If we write this in matrix form, so that:
\begin{equation}
A_i (r^2 \dw)_{i-\thalf} + B_i (r^2 \dw)_{i-\half} + C_i (r^2 \dw)_{i+\half} = F_i\enskip,
\end{equation}
then:
\begin{eqnarray}
A_i &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{i-1}  \enskip, \nonumber  \\
B_i &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar p_0}{r^2}\right)_{i  }  
               +\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \right] 
              - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} \nonumber \\
C_i &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{i}  \enskip, \nonumber  \\
F_i &=&  \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
 - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip \nonumber .
\end{eqnarray}
Here,
\begin{equation}
g_{j-\half} = \frac{G}{r_{j-\half}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\half}^3 - r_{k-\thalf}^3) \; {\rho_0}_{k-1} \enskip.
\end{equation}

We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
corresponds to $i=1,$ by setting:
\begin{eqnarray}
A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
B_1 &=& 1\enskip. \nonumber
\end{eqnarray}
We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $i=N,$
by setting:
\begin{eqnarray}
A_N &=& -1\enskip, \nonumber \\
B_N &=&  1\enskip, \nonumber \\
C_N = F_N &=& 0\enskip. \nonumber
\end{eqnarray}
\end{description}

@@

\item[plane-parallel:] We discretize the base state density evolution 
equation from paper III equation (29), without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ]
         \nonumber \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are found via the
procedure described in paper II, Appendix A.\\

@@

First we compute a cell-centered $\psi^\outp$ using
\begin{equation}
 \psi_j^\outp = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
\end{equation}
Then, we discretize the base state pressure evolution equation from paper III equation (31),
\begin{equation}
\frac{\partial p_0}{\partial t} = 
-w_0 \frac{\partial p_0}{\partial r} + \psi \enskip ,
\label{eq:p0upd_new}
\end{equation}
to compute the new base state pressure,
\begin{equation}
p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\outp 
\enskip,
\end{equation}
where the interface states are again found via the procedure described
in paper II, Appendix A.\\

@@

\item At the beginning of each time step, we need an estimate for the time-centered
source term in the velocity divergence constraint, as given in paper III equation (19),
\begin{equation}
  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext 
  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip .
\label{eq:defineS} 
\end{equation}

@@

\item {\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,\psi^{n-\myhalf}]
\rightarrow [w_0^{n+\myhalf,*}]$

@@

The local velocity field is described by paper III equation (37),
\begin{equation}
\frac{\partial\Ubt}{\partial t} = - \Ubt \cdotb \nablab \Ubt - w_0 \frac{\partial \Ubt}{\partial r}
                                 - \left(\Ubt \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                 - \frac{1}{\rho} \nablab\pi
                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
 .
\end{equation}
